---
title: "p8105_hw3_yl4923"
output: github_document
---

```{r setup, include=FALSE}
library(tidyverse)
library(p8105.datasets)
data("instacart")
data("brfss_smart2010")

library(readxl)
knitr::opts_chunk$set(
  echo = TRUE,
  fig.width = 10,
  fig.asp = .6,
  out.width = "90%")

theme_set(theme_bw() + theme(legend.position = "bottom"))
```

## Problem 1
1.1 data exploration
There are *`r nrow(instacart)`* rows and *`r ncol(instacart)`* columns in the _instacart_ data set. The key variables like *order_id*, *product_id* or *user_id* helps both customers and stores to identify products bought. For instant, in row 1, *user 112108* ordered *Bulgarian Yogurt (product 49302)* which was the first to be added to the cart and was also reordered by this user. This product was placed at *yogurt aisle (aisle_id 120)* in the *dairy & eggs department (department_id 16)*.The order was placed on Thursday between 10-11 am, 9 days after the user's last purchase. 

1.2 aisles
```{r count aisle}
instacart %>% 
  group_by(aisle, aisle_id) %>%
  summarize(n = n()) %>%
  arrange(desc(n)) %>%
  tbl_df()
```
There are *134* different aisles and most items ordered with a count of 150609 is from aisle *fresh vegetables* with id _83_. 

1.3 plot aisles with more 10000 items ordered
```{r plot}
instacart %>% 
  group_by(aisle, department) %>% 
  summarise(n = n()) %>%
  filter(n > 10000) %>% 
  ggplot(aes(x = reorder(aisle, n), y = n, fill = department)) + 
  geom_histogram(position = "dodge", stat = "identity") +
  labs(
    title = "Aisles with more than 10k ordered",
    x = "Aisle",
    y = "Number of items ordered",
    caption = "Instacart data") + 
   theme(axis.text.x = element_text(angle = 90, vjust = 0.25, hjust = 1),
        axis.text = element_text(size = 7.5) )
```

  The histogram shows the number of products in aisles with more than 10k orders as well as colored them which department falls in. Thus, we can tell these aisles by more general categories, which can be easier to find than through their long names.  
  
1.4 a table showing the three most popular items in each of the aisles “baking ingredients”, “dog food care”, and “packaged vegetables fruits”
```{r table 1}
instacart %>%
  filter(aisle == c("baking ingredients", "dog food care", "packaged vegetables fruits")) %>%
  group_by(aisle, product_name) %>%
  summarize(order_number = n()) %>% 
  filter(order_number == max(order_number)) %>%
  knitr::kable(caption = "Most popular items of the aisles “baking ingredients”, “dog food care”, and “packaged vegetables fruits” ", align = "lccrr")
```
The 3 most popular items are *Light Brown Sugar*, *Organic Grain Free Chicken & Vegetable Dog Food* and *Organic Baby Spinach* with 156, 14 and 3324 times of order respectively.

1.5 a table showing the mean hour of the day at which Pink Lady Apples and Coffee Ice Cream are ordered on each day of the week
```{r table 2}
instacart %>%
  filter(product_name %in% c("Pink Lady Apples", "Coffee Ice Cream")) %>%
  select(product_name, order_dow, order_hour_of_day) %>% 
  group_by(product_name, order_dow) %>% 
  summarize(mean_hour = mean(order_hour_of_day)) %>% 
  mutate(order_dow_char = c("Sun", "Mon", "Tue", "Wed", "Thu", "Fri", "Sat")) %>%
  select(-order_dow) %>% 
  pivot_wider(
    names_from = "order_dow_char", 
    values_from = "mean_hour") %>% 
  knitr::kable(digits = 2,
               caption = "Mean hour of each day at which Pink Lady Apples and Coffee Ice Cream are ordered", 
               align = "lccrr")
```
Pink Lady Apple is mostly ordered around noon on each day of the week, ranging from 11.6 to 13.9 while Coffee Ice Cream is more likely to be ordered in the afternoon, ranging from 12.4 to 15.4 with one exception in the morning at 10.3 by average on Friday.

### Problem 2
2.1 Data cleaning 
```{r data cleaning}
brfss_df =
  brfss_smart2010 %>%
  janitor::clean_names() %>%
  filter(topic == "Overall Health") %>% 
  mutate(response = fct_relevel(response, c("Poor", "Fair", "Good", "Very good", "Excellent"))) %>% 
  select(year,
         state = locationabbr,
         location = locationdesc,
         response,
         sample_size,
         data_value)
```

2.2 States observed at 7 or more locations in 2002 and 2010
```{r}
state_2002 = 
  brfss_df %>% 
  filter(year == 2002) %>% 
  distinct(state, location) %>%
  count(state) %>%
  filter(n >= 7) %>%
  tbl_df

state_2010 = 
  brfss_df %>% 
  filter(year == 2010) %>% 
  distinct(state, location) %>%
  count(state) %>%
  filter(n >= 7) %>%
  tbl_df
```
In 2002,*`r pull(state_2002, state)`* were observed at 7 or more locations; in 2010, *`r pull(state_2010, state)`* were observed at 7 or more locations.

2.2 Plot 'Excellent' response
```{r}
brfss_df %>% 
  filter(response == "Excellent") %>% 
  select(year, state, data_value) %>% 
  group_by(year, state) %>% 
  summarize(ave_data = mean(data_value)) %>%
  ggplot(aes(x = year, y = ave_data, color = state)) + 
  geom_line() +
  labs(
    title = "Spaghetti plot of average values across locations within a state from 2002 to 2010",
    y = "average data value",
    caption = "BRFSS 'Excellent' response for overall health")
```
Most of states appears a relatively constant average *data value* over the eight year period while state AK in red line is the one deviates the most in 2005 and 2009.


2.3 Plot 2006 and 2010 health responses in NY 
```{r two-panel plot}
brfss_df %>% 
  filter(state == "NY", year == c("2006", "2010")) %>%
  select(year, location, response, data_value) %>% 
  ggplot(aes(x = response, y = data_value, color = location)) + 
  geom_point() + 
  facet_grid(. ~ year) +
  labs(
    title = "Overall Health Response in NY",
    x = "Health Response",
    y = "Distribution of Data Value",
    caption = "BRFSS IN NY")
```
The response to overall health topic seems getting better compared 2006 to 2010 in New York State since "Poor" and "Fair" lessen while "Good" and "Very Good" increases. The distribution for answer of "Excellent" is about the same.  

### Problem 3 
3.1 data cleaning
```{r import and tidy data }
acc_df = read_csv("data/accel_data.csv") %>% 
  janitor::clean_names() %>%
  pivot_longer(
    activity_1:activity_1440,
    names_to = "activity", 
    names_prefix = "activity_",
    values_to = "n_act"
  ) %>% 
  mutate(day_class = recode(day,
                            "Monday" = "weekday", "Tuesday" = "weekday", "Wednesday" = "weekday", "Thursday" = "weekday", "Friday" = "weekday", "Saturday" = "weekend", "Sunday" = "weekend"))
         
```
The resulting data set includes, _week_, _day_id_, _day_ as well as _activity_ and _n_act_ derived from original _activity_1_ to _activity_1440_. Also, a new variable _day_class_ classified a _day_ into weekday or weekend is added. We now have *`r nrow(acc_df)`* observations and *`r ncol(acc_df)`* variables for this five weeks of accelerometer data. 


3.2 total activity table for each day
```{r daily activity table}
acc_df %>%
  group_by(week, day) %>% 
  summarize(total_act = sum(n_act)) %>%
  mutate(day = c("Monday", "Tuesday", "Wednesday", "Thursday","Friday", "Saturday", "Sunday")) %>%
  select(week, day, total_act) %>% 
  knitr::kable(caption = "Total physical activity of each day",
               align = "lccrr")
```
For the first 3 weeks, participants seems more active on Monday and midweek (Wednesday or Thursday). Fourth week has a noticeable drop in total activity for each day while the last week participants regained similar amount of activity and pattern as the first 3 weeks to be more active on Monday and Friday than other days of the week.

3.3 24-hour activity time courses for each day
```{r 24h activity}
acc_df %>%
  group_by(week, day) %>% 
  ggplot(aes(x = activity, y = n_act, color = day)) + 
  geom_line() +
  labs(
    title = "Daily Activity Over 5 Weeks",
    x = "Day of Week",
    y = "Total Activity",
    caption = "Accelerometer Data")
```
Inspecting the activity over each day through 5 weeks, it seems like the participants are more fond of exercising during weekdays at the beginning of the study in which more weekdays(more organge yellow or pink then greens) have high physical activity peaks in the first third proportion of the graph. For the left weeks, there is a obvious gap in the mid of study while then the participants seem more into doing physical activity on weekend as more peaks showed on Saturday or Sunday (more greens) near the end of study.

