---
title: "p8105_hw3_yl4923"
output: github_document
---
### Problem 0 

```{r setup, include=FALSE}
library(tidyverse)
library(p8105.datasets)
data("instacart")
data("brfss_smart2010")

library(readxl)
knitr::opts_chunk$set(
  echo = TRUE,
  fig.width = 10,
  fig.asp = .6,
  out.width = "90%")

theme_set(theme_bw() + theme(legend.position = "bottom"))
```

## Problem 1
1.1 data exploration
There are *`r nrow(instacart)`* rows and *`r ncol(instacart)`* columns in the _instacart_ data set. The key variables like *order_id*, *product_id* or *user_id* helps both customers and stores to identify products bought. For instant, in row 1, *user 112108* ordered *Bulgarian Yogurt (product 49302)* which was the first to be added to the cart and was also reordered by this user. This product was placed at *yogurt aisle (aisle_id 120)* in the *dairy & eggs department (department_id 16)*.The order was placed on Thursday between 10-11 am, 9 days after the user's last purchase. 

1.2 aisles
```{r count aisle}
instacart %>% 
  group_by(aisle, aisle_id) %>%
  summarize(n = n()) %>%
  arrange(desc(n)) %>%
  tbl_df()
```
There are *134* different aisles and most items ordered with a count of 150609 is from aisle *fresh vegetables* with id _83_. 